Smoking and COVID Living Review (v10): a bayesian analysis

### 2021-01-03 reports/bayesian_report.Rmd
Update for version 10, change to plots

---
title: 'Smoking and COVID Living Review (v8): a bayesian analysis'
author: "David Simons, Lion Shahab, Jamie Brown and Olga Perski"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:

@@

## Load packages
* **Load packages**: load packages and useful scripts for the analysis
* **Load scripts**: We have written functions to handle much of the Bayesian analysis
* **Load Bayesian datasets**: The Bayesian approach relies on large simulated datasets from the observed data, for this report we will only present the results graphically from data we have simulated elsewhere. The scripts `bayes_scripts.R`, `bayesian_data_generation.R` and `bayesian_analysis.R` should be able to be run locally if you want to generate your own simulated data. Alternatively you can use the pre-generated data which is stored in the `data_clean\bayesian_models` folder in the github repository [here](https://github.com/DidDrog11/smoking_covid/tree/master) and then you can skip the `bayesian_data_generation.R` step.
```{r install_packages, include=FALSE}
library(brms)
library(ggplot2)
library(tidybayes)
library(readr)
library(dplyr)
library(tidyr)
library(ggridges)
library(glue)
library(stringr)
library(forcats)
library(here)
library(meta)
library(snakecase)
library(flextable)
library(googlesheets4)
library(linelist)
library(DT)
library(epitools)
library(ggmap)
```

```{r load_scripts}
source(here("scripts", "author_dictionary.R"))
```
## Download data
``` {r google_data}
sheets_id <- as_sheets_id("https://docs.google.com/spreadsheets/d/1yLtpbk29oUKOGceVw2IAJlt9CbbbPWmltzOW_LG6Su8/edit?usp=sharing")
search_details <- read_sheet(sheets_id, range = 'article_screening')
data_study_general <- read_sheet(sheets_id, range =  'general_details')
table_1 <- read_sheet(sheets_id, sheet = 'pop_descriptives')
table_2 <-  read_sheet(sheets_id, sheet = 'testing')
table_3 <-  read_sheet(sheets_id, sheet = 'hospitalisation')
table_4 <-  read_sheet(sheets_id, sheet = 'severity')
table_5 <-  read_sheet(sheets_id, sheet = 'mortality')
table_6 <-  read_sheet(sheets_id, sheet = 'updated_quality_appraisal')
```
We maintain a Google sheets document [here](sheets_id) which contains all eligible articles the research team have identified. This is currently viewable only and not open for editing. We have collected the general information for the search in the first sheet, the studies general information in the second, further information on study populations, outcomes of interest and quality of the studies are contained in the further sheets. These sheets are organised in a similar way to the manuscript and so are labelled table_1 to table_6.

@@

```{r head_search_details}
flextable(search_details[,1:12])
```
The second sheet is **General study details** this includes the lead author, where the study is published, the date it was published, the country in which the data originated and the DoI or link to the publication for our records.
This is a searchable list of all the studies that have met out inclusion and exclusion criteria up until `r paste(max(search_details$date_screened))`. These are live links and should direct to the relevant article, however, some may have been deleted or require institutional access.
```{r tail_study_details, include=FALSE}
datatable(data_study_general[1:4] %>%
            mutate(`date_published_format(2020-05-12)` = as.Date(data_study_general$`date_published_format(2020-05-12)`),
                   doi = paste0("<a href='",doi,"' target='_blank'>",doi,"</a>")),
          colnames = c("Lead author" = 2, "Date published" = 3, "Source" = 4, "DOI" = 5),
          escape = F)
```
The third sheet produces **Table 1** this contains the details of the population of the studies in our review. Primarily includes smoking status but also the proportion of females in the study and the age of participants.
```{r tail_table_1, include=FALSE}
datatable(table_1 %>%
                 filter(lead_author != is.na(lead_author)))
```
The fourth sheet produces **Table 2** this contains the data extracted from relevant studies on testing data and smoking status. Most studies will not have data on this outcome however we filter these subsequently. These are sparse datasets as many studies do not report on all levels of smoking exposure.
```{r tail_table_2, include=FALSE}
datatable(
  table_2 %>%
  filter(data_on_testing == TRUE)
)
```
The fifth sheet produces **Table 3** this contains the data extracted from relevant studies on hospitalisation and smoking status. 
```{r tail_table_3, include=FALSE}
datatable(
  table_3 %>%
  filter(data_on_hospitalisation == TRUE)
)
```
The sixth sheet produces **Table 4** this contains the data extracted from relevant studies disease severity data and smoking status. Disease severity is differentially reported across the studies in some there is reference to WHO severity score in others we have based the distinction on clinical characteristics or whether patients were admitted to Intensive or Critical care areas.
```{r tail_table_4, include=FALSE}
datatable(
  table_4 %>%
  filter(data_disease_severity == TRUE)
)
```
The seventh sheet produces **Table 5** this contains the data extracted from relevant studies on COVID-19 related mortality and smoking status. A caution needed when interpreting this data is that most studies have not followed patients up for sufficient amounts of time to be able to accurately report on disease associated mortality.
```{r tail_table_5, include=FALSE}
datatable(
  table_5 %>%
  filter(data_on_deaths == TRUE)
)
```
The final sheet is used to collate data on **Study quality** we changed our method on assessing study quality in version 3 of the review. We use the following criteria:

@@

table_1 <- table_1 %>%
  clean_data() %>%
  mutate(., lower_range = sub('\\_.*', '', .$range)) %>%
  mutate(., upper_range = sub('.*_', '',.$range,)) %>% ## We break apart the age range into an upper and lower level
  select(lead_author, sample_size, female_sex_percent, data_source, median_age, mean_age, iqr_lower, iqr_upper, standard_deviation, lower_range,
         upper_range, current_smoker, former_smoker, current_former_smoker, never_smoker, never_smoker_unknown, current_vaper, former_vaper,
         current_former_vaper, not_stated, missing, total) %>%
  filter(!is.na(lead_author)) %>%
  left_join(., review_details, by = 'lead_author') %>%
  write_rds(here::here('data_clean', 'table_1.rds'))

@@

table_2 <-  table_2 %>%
  clean_data() %>%
  filter(data_on_testing == TRUE) %>%
  left_join(., review_details, by = 'lead_author') %>%
  write_rds(here::here('data_clean', 'table_2.rds'))

table_3 <-  table_3 %>%
  clean_data() %>%
  filter(data_on_hospitalisation == TRUE) %>%
  left_join(., review_details, by = 'lead_author') %>%
  write_rds(here::here('data_clean', 'table_3.rds'))

table_4 <-  table_4 %>%
  clean_data() %>%
  filter(data_disease_severity == TRUE) %>%
  left_join(., review_details, by = 'lead_author') %>%
  write_rds(here::here('data_clean', 'table_4.rds'))

table_5 <- table_5 %>%
  clean_data() %>%
  filter(data_on_deaths == TRUE) %>%
  left_join(., review_details, by = 'lead_author') %>%
  write_rds(here::here('data_clean', 'table_5.rds'))

@@

table_6 <-  table_6 %>%
  clean_data(protect = protect_columns_2) %>%
  left_join(., table_1 %>%
              select(lead_author,
                     not_stated,
                     missing,
                     total),
                          by = 'lead_author') %>%
  mutate(missingness = rowSums(.[8:9], na.rm = T),
         missingness_percentage = (missingness/total)*100) %>% ## Later we want to assess the amount of missingness for each study as it would impact the inclusion of a study into the relevant meta-analyses
  select(-c(not_stated, missing, total, missingness)) %>%
  left_join(., review_details, by = 'lead_author') %>%
  write_rds(here::here('data_clean', 'table_6.rds'))

@@

exclude_from_analysis <- c(superseded_studies, old_isaric, excluded_other_reason)
```
We don't want to remove studies from the dataset despite them no longer being included in the analysis. The excluded studies are defined in this chunk of code. We have excluded this version the following studies:

@@

## The PRISMA diagram will be updated
The PRISMA reporting guidelines for systematic reviews the guidance is available [here](http://www.prisma-statement.org/). This takes the format of the number of articles obtained from the search from [OVID](https://ovidsp.dc1.ovid.com/sp-4.07.0b/ovidweb.cgi) (login may be needed), [medRxiv](https://www.medrxiv.org/search/text_abstract_title%3Asmoking%2BAND%2BCOVID%20text_abstract_title_flags%3Amatch-all%20jcode%3Amedrxiv%20numresults%3A75%20sort%3Apublication-date%20direction%3Adescending%20format_result%3Astandard), and articles identified by the research team. These are then screened at the title and abstract level. Full texts are then read to extract relevant data, if articles are excluded at this level we describe the reason for this. All of the selected articles will contribute to the narrative synthesis but only those that are judged 'good' or 'fair' quality will be included in the Bayesian meta-analysis.
``` {r PRISMA}
current_total <- search_details %>%
  select(total_included) %>%
  sum()

previous_total <- search_details %>%
  filter(review_version %in% prev_versions) %>%
  select(total_included) %>%
  sum()

prisma <- search_details %>%
  group_by(review_version) %>%
  filter(review_version == current_version) %>%
  mutate(title_abstract = sum(ovid_number_results, medrxiv_number_results, other_source_included),
         full_texts_assessed = sum(ovid_screened, medrxiv_screened, other_source_included),
         title_abstract_included = sum(ovid_included, medrxiv_included, other_source_included),
         title_abstract_excluded = title_abstract - full_texts_assessed,
         full_text_excluded = full_texts_assessed-total_included,
         previous_total = previous_total,
         previous_minus_superseded = previous_total-superseded,
         current_total = current_total-superseded) %>%
  select("Review version" = review_version,
         "Date Screened" = date_screened,
         "OVID results" = "ovid_number_results",
         "medRxiv results" = "medrxiv_number_results",
         "Other sources" = "other_source_included",
         "Title/abstracts screened" = "title_abstract",
         "Title/abstracts excluded" = "title_abstract_excluded",
         "Full texts assessed" = "full_texts_assessed",
         "Full texts excluded" = "full_text_excluded",
         "Records carried forward" = "previous_minus_superseded",
         "Supersed from previous version" = superseded,
         "Studies in narrative synthesis" = "current_total")
flextable(prisma)
```
The number of studies contributing to meta-analysis is calculated later.

@@

``` {r countries, cache = TRUE}
library('ggmap')
#Countries
country <- table_1$country %>%
  to_upper_camel_case() %>%
  recode(., 'Usa' = 'USA', 'Uk' = 'UK', 'SaudiArabia' = 'Saudi Arabia', "SouthKorea" = "South Korea") %>%
  table() %>%
  as.data.frame() %>%
  arrange(desc(Freq)) %>%
  rename('Country' = 1,
         'Number' = 2) %>%
  mutate(Country = as.character(Country))

@@

my_bins <- c(0, 1, 5, 10, 20, 30, 40, 80)
mypalette <- colorBin(palette="YlOrBr", domain=my_spdf@data$Number, na.color="transparent", bins=my_bins)

@@

* **Summary statistics:** this table will give us the mean, median and IQR of the samples in the review 
``` {r numbers}
#Numbers
summary(table_1$sample_size)
```
* **Total number of participants**: `r paste(sum(table_1$sample_size))`

@@

b$`Lead author` <- to_upper_camel_case(b$`Lead author`, sep_out = ", ")
b$`Lead author` <- plyr::mapvalues(b$`Lead author`,
                               from = cleaned_names,
                               to = correct_names)  
datatable(b,
          caption = "Studies providing data on current, former and never smokers") %>%
  formatRound(columns = c(2:5), digits = 0, interval = 3, mark = ",")
#These are the studies with data on current, former and never smokers

b <- b %>%
  select(-"Study ID") %>%
  flextable() %>%
    set_caption(caption = 'Studies reporting complete smoking status') %>%
  colformat_num(col_keys = c("Sample size", "Current smokers", "Former smokers", "Never smokers"),
                digits = 0, na_str = '-', big.mark = ',') %>%
  set_table_properties(width = 1, layout = 'autofit')

@@

b <- b %>%
  select(-"Study ID") %>%
  flextable() %>%
    set_caption(caption = 'Studies reporting complete smoking status') %>%
  colformat_num(col_keys = c("Sample size", "Current/former smokers", "Never smokers"),
                digits = 0, na_str = '-', big.mark = ',') %>%
  set_table_properties(width = 1, layout = 'autofit')

@@

numeric_names <- colnames(b[2:9])
b <- b %>%
  select(-"Study ID") %>%
  flextable() %>%
    set_caption(caption = 'Studies reporting complete smoking status') %>%
  colformat_num(col_keys = numeric_names,
                digits = 0, na_str = '-', big.mark = ',') %>%
  set_table_properties(width = 1, layout = 'autofit')

@@

# Smoking prevalence by country {.tabset .tabset-fade .tabset-pills}
``` {r smoking_country}

@@

```
We have split the data into a list where each country can be analysed separately if required.

@@

a <- flextable(a) %>%
    set_caption(caption = 'Characteristics of included studies') %>%
  colformat_num(col_keys = numeric_columns, digits = 1, na_str = '-', big.mark = ',') %>%
  colformat_num(col_keys = 'Sample size', digits = 0, na_str = '-', big.mark = ',') %>%
  set_table_properties(width = 1, layout = 'autofit')

@@

numeric_columns <- c('Population_tested', 'SARS_CoV_2_negative', 'N_current_smoker', 'N_former_smoker',
                     'N_current_former_smoker', 'N_never_smoker', 'N_not_stated', 'SARS_CoV_2_positive', 'P_current_smoker',
                     'P_former_smoker', 'P_current_former_smoker', 'P_never_smoker', 'P_not_stated')
a <- flextable(a) %>%
    set_caption(caption = 'SARS-CoV-2 infection by smoking status') %>%
  colformat_num(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',')
```
The number of studies reporting on infection by smoking status is: `r paste(nrow(table_2_word))`
``` {r}
a <- set_header_labels(a,
                       Population_tested = 'Total population tested',
                       SARS_CoV_2_negative = "N (%)",
                       N_current_smoker = "Current smoker (%)",
                       N_former_smoker = "Former smoker (%)",
                       N_current_former_smoker = "Current/former smoker (%)",
                       N_never_smoker = "Never smoker (%)",
                       N_not_stated = "Not stated (%)",
                       SARS_CoV_2_positive = "N (%)",
                       P_current_smoker = "Current smoker (%)",
                       P_former_smoker = "Former smoker (%)",
                       P_current_former_smoker = "Current/former smoker (%)",
                       P_never_smoker = "Never smoker (%)",
                       P_not_stated = "Not stated (%)") %>%
  add_header_row(top = TRUE, values = c("","SARS-CoV-2 negative", "SARS-CoV-2 positive" ), colwidths = c(2, 6, 6)) %>%
  theme_booktabs() %>%
  fix_border_issues() %>%
  set_table_properties(width = 1, layout = 'autofit')  %>%
  colformat_char(na_str = '-')
a

@@

a <- flextable(a) %>%
    set_caption(caption = 'COVID-19 hospitalisation by smoking status') %>%
  colformat_num(col_keys = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
  colformat_char(na_str = '-')

@@

a <- flextable(a) %>%
    set_caption(caption = 'COVID-19 severity by smoking status') %>%
  colformat_num(col_keys = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
  colformat_char(na_str = '-')

@@

a <- flextable(a) %>%
    set_caption(caption = 'COVID-19 mortality by smoking status') %>%
  colformat_num(col_keys = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
  colformat_char(na_str = '-')

@@

country_smoking <-  read_sheet(sheets_id, sheet = 'national_smoking_prevalence') %>%
  flextable()
save_as_docx(country_smoking, path = here('output_data', 'Supplementary_3.docx'))
```

>>>>

---
title: 'Smoking and COVID Living Review (v10): a bayesian analysis'
author: "David Simons, Lion Shahab, Jamie Brown and Olga Perski"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:

@@

## Load packages
* **Load packages**: load packages and useful scripts for the analysis
* **Load scripts**: We have written functions to handle much of the Bayesian analysis
* **Load Bayesian datasets**: The Bayesian approach relies on large simulated datasets from the observed data, for this report we will only present the results graphically from data we have simulated elsewhere. The scripts `bayes_scripts.R`, `bayesian_data_generation.R` and `bayesian_analysis.R` should be able to be run locally if you want to generate your own simulated data. Alternatively you can use the pre-generated data which is stored in the `data_clean\bayesian_models` folder in the github repository [here](https://github.com/DidDrog11/smoking_covid/tree/master) and then you can skip the `bayesian_data_generation.R` step.
```{r load_packages, include=FALSE, quietly=T}
source(here::here("scripts", "libraries.R"))
```

```{r load_dictionary}
source(here("scripts", "author_dictionary.R"))
```
## Download data
``` {r google_data}
google_api <- rstudioapi::askForSecret("Google API Key")
register_google(google_api)
drive_download("https://docs.google.com/spreadsheets/d/1jelIGxrm99QpDKwSoyMUliDJQWtzk4bHsNGDaIjxTaU/edit?usp=sharing", path = here("data_spreadsheet", "data_extraction_current.xlsx"), overwrite = T)

search_details <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'article_screening')
data_study_general <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'general_details')
table_1 <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'pop_descriptives')
table_2 <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'testing')
table_3 <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'hospitalisation')
table_4 <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'severity')
table_5 <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'mortality')
table_6 <- readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'updated_quality_appraisal')
```
We maintain a Google sheets document [here](sheets_id) which contains all eligible articles the research team have identified. This is currently viewable only and not open for editing. We have collected the general information for the search in the first sheet, the studies general information in the second, further information on study populations, outcomes of interest and quality of the studies are contained in the further sheets. These sheets are organised in a similar way to the manuscript and so are labelled table_1 to table_6.

@@

```{r head_search_details}
flextable(search_details[,1:13])
```
The second sheet is **General study details** this includes the lead author, where the study is published, the date it was published, the country in which the data originated and the DoI or link to the publication for our records.
This is a searchable list of all the studies that have met out inclusion and exclusion criteria up until `r paste(max(search_details$date_screened))`. These are live links and should direct to the relevant article, however, some may have been deleted or require institutional access.
```{r tail_study_details, include=FALSE}
datatable(data_study_general[1:4] %>%
            mutate(`date_published_format(2020-05-12)` = as.Date(data_study_general$`date_published_format(2020-05-12)`),
                   doi = paste0("<a href='",doi,"' target='_blank'>",doi,"</a>")),
          colnames = c("Lead author" = 2, "Date published" = 3, "Source" = 4, "DOI" = 5),
          escape = F)
```
The third sheet produces **Table 1** this contains the details of the population of the studies in our review. Primarily includes smoking status but also the proportion of females in the study and the age of participants.
```{r tail_table_1, include=FALSE}
datatable(table_1 %>%
                 filter(lead_author != is.na(lead_author)))
```
The fourth sheet produces **Table 2** this contains the data extracted from relevant studies on testing data and smoking status. Most studies will not have data on this outcome however we filter these subsequently. These are sparse datasets as many studies do not report on all levels of smoking exposure.
```{r tail_table_2, include=FALSE}
datatable(
  table_2 %>%
  filter(data_on_testing == TRUE)
)
```
The fifth sheet produces **Table 3** this contains the data extracted from relevant studies on hospitalisation and smoking status. 
```{r tail_table_3, include=FALSE}
datatable(
  table_3 %>%
  filter(data_on_hospitalisation == TRUE)
)
```
The sixth sheet produces **Table 4** this contains the data extracted from relevant studies disease severity data and smoking status. Disease severity is differentially reported across the studies in some there is reference to WHO severity score in others we have based the distinction on clinical characteristics or whether patients were admitted to Intensive or Critical care areas.
```{r tail_table_4, include=FALSE}
datatable(
  table_4 %>%
  filter(data_disease_severity == TRUE)
)
```
The seventh sheet produces **Table 5** this contains the data extracted from relevant studies on COVID-19 related mortality and smoking status. A caution needed when interpreting this data is that most studies have not followed patients up for sufficient amounts of time to be able to accurately report on disease associated mortality.
```{r tail_table_5, include=FALSE}
datatable(
  table_5 %>%
  filter(data_on_deaths == TRUE)
)
```
The final sheet is used to collate data on **Study quality** we changed our method on assessing study quality in version 3 of the review. We use the following criteria:

@@

table_1 <- table_1 %>%
  clean_data() %>%
  mutate(., lower_range = sub('\\_.*', '', .$range)) %>%
  mutate(., upper_range = sub('.*_', '',.$range,)) %>% ## We break apart the age range into an upper and lower level
  select(lead_author, sample_size, female_sex_percent, data_source, median_age, mean_age, iqr_lower, iqr_upper, standard_deviation, lower_range,
         upper_range, current_smoker, former_smoker, current_former_smoker, never_smoker, never_smoker_unknown, current_vaper, former_vaper,
         current_former_vaper, not_stated, missing, total) %>%
  mutate(study_id = 1:length(lead_author)) %>%
  filter(!is.na(lead_author)) %>%
  left_join(., review_details %>%
              select(-"lead_author"), by = 'study_id') %>%
  write_rds(here::here('data_clean', 'table_1.rds'))

@@

table_2 <-  table_2 %>%
  clean_data() %>%
  mutate(study_id = 1:length(lead_author)) %>%
  filter(data_on_testing == TRUE) %>%
  left_join(., review_details %>%
              select(-"lead_author"), by = 'study_id') %>%
  write_rds(here::here('data_clean', 'table_2.rds'))

table_3 <-  table_3 %>%
  clean_data() %>%
  mutate(study_id = 1:length(lead_author)) %>%
  filter(data_on_hospitalisation == TRUE) %>%
  left_join(., review_details %>%
              select(-"lead_author"), by = 'study_id') %>%
  write_rds(here::here('data_clean', 'table_3.rds'))

table_4 <-  table_4 %>%
  clean_data() %>%
  mutate(study_id = 1:length(lead_author)) %>%
  filter(data_disease_severity == TRUE) %>%
  left_join(., review_details %>%
              select(-"lead_author"), by = 'study_id') %>%
  write_rds(here::here('data_clean', 'table_4.rds'))

table_5 <- table_5 %>%
  clean_data() %>%
  mutate(study_id = 1:length(lead_author)) %>%
  filter(data_on_deaths == TRUE) %>%
  left_join(., review_details %>%
              select(-"lead_author"), by = 'study_id') %>%
  write_rds(here::here('data_clean', 'table_5.rds'))

@@

table_6 <-  table_6 %>%
  clean_data(protect = protect_columns_2) %>%
  mutate(study_id = 1:length(lead_author)) %>%
  left_join(., table_1 %>%
              select(not_stated,
                     missing,
                     total,
                     study_id),
                          by = 'study_id') %>%
  mutate(missingness = rowSums(.[8:9], na.rm = T),
         missingness_percentage = (missingness/total)*100) %>% ## Later we want to assess the amount of missingness for each study as it would impact the inclusion of a study into the relevant meta-analyses
  select(-c(not_stated, missing, total, missingness)) %>%
  left_join(., review_details %>%
              select(-lead_author), by = 'study_id') %>%
  write_rds(here::here('data_clean', 'table_6.rds'))

@@

exclude_from_analysis <- unique(c(superseded_studies, old_isaric, excluded_other_reason))
write_rds(exclude_from_analysis, here("output_data", "excluded_studies.rds"))
```
We don't want to remove studies from the dataset despite them no longer being included in the analysis. The excluded studies are defined in this chunk of code. We have excluded this version the following studies:

@@

## The PRISMA diagram will be updated
The PRISMA reporting guidelines for systematic reviews the guidance is available [here](http://www.prisma-statement.org/). This takes the format of the number of articles obtained from the search from [OVID](https://ovidsp.dc1.ovid.com/sp-4.07.0b/ovidweb.cgi) (login may be needed), [medRxiv](https://www.medrxiv.org/search/text_abstract_title%3Asmoking%2BAND%2BCOVID%20text_abstract_title_flags%3Amatch-all%20jcode%3Amedrxiv%20numresults%3A75%20sort%3Apublication-date%20direction%3Adescending%20format_result%3Astandard), and articles identified by the research team. These are then screened at the title and abstract level. Full texts are then read to extract relevant data, if articles are excluded at this level we describe the reason for this. All of the selected articles will contribute to the narrative synthesis but only those that are judged 'good' or 'fair' quality will be included in the Bayesian meta-analysis.
``` {r PRISMA}
current_total <- 345

previous_total <- 279

prisma <- search_details %>%
  group_by(review_version) %>%
  filter(review_version == current_version) %>%
  mutate(title_abstract = sum(ovid_titles, medrxiv_titles, other_source_included),
         full_texts_assessed = sum(ovid_abstract_full, medrxiv_abstract_full, other_source_included),
         title_abstract_included = sum(ovid_included, medrxiv_included, other_source_included),
         title_abstract_excluded = title_abstract - full_texts_assessed,
         full_text_excluded = full_texts_assessed-total_included,
         full_text_included = full_texts_assessed-full_text_excluded,
         previous_total = previous_total,
         previous_minus_superseded = previous_total-superseded,
         current_total = current_total) %>%
  select("Review version" = review_version,
         "Date Screened" = date_screened,
         "OVID results" = "ovid_titles",
         "medRxiv results" = "medrxiv_titles",
         "Other sources" = "other_source_included",
         "Titles/abstracts screened" = "title_abstract",
         "Titles/abstracts excluded" = "title_abstract_excluded",
         "Full texts screened" = "full_texts_assessed",
         "Full texts excluded" = "full_text_excluded",
         "Records included" = "full_text_included",
         "Records carried forward" = "previous_minus_superseded",
         "Supersed from previous version" = superseded,
         "Studies in narrative synthesis" = "current_total")
flextable(prisma)
```
The number of studies contributing to meta-analysis is calculated later.

@@

``` {r countries, cache = TRUE}
#Countries
country <- table_1$country %>%
  to_upper_camel_case() %>%
  recode(., 'Usa' = 'USA', 'Uk' = 'UK', 'SaudiArabia' = 'Saudi Arabia', "SouthKorea" = "South Korea", "UnitedArabEmirates" = "United Arab Emirates") %>%
  table() %>%
  as.data.frame() %>%
  arrange(desc(Freq)) %>%
  rename('Country' = 1,
         'Number' = 2) %>%
  mutate(Country = as.character(Country))

@@

my_bins <- c(0, 1, 5, 10, 20, 50, 70, 100)
mypalette <- colorBin(palette="YlOrBr", domain=my_spdf@data$Number, na.color="transparent", bins=my_bins)

@@

* **Summary statistics:** this table will give us the mean, median and IQR of the samples in the review 
``` {r numbers}
#Numbers
summary(as.numeric(table_1$sample_size))
```
* **Total number of participants**: `r paste(sum(as.numeric(table_1$sample_size)))`

@@

b$`Lead author` <- to_upper_camel_case(b$`Lead author`, sep_out = ", ")
b$`Lead author` <- plyr::mapvalues(b$`Lead author`,
                               from = cleaned_names,
                               to = correct_names)  
datatable(b,
          caption = "Studies providing data on current, former and never smokers") %>%
  formatRound(columns = c(2:5), digits = 0, interval = 3, mark = ",")
#These are the studies with data on current, former and never smokers
set_flextable_defaults(digits = 0, big.mark = ",", na_str = "-")

b <- b %>%
  select(-"Study ID") %>%
  flextable() %>%
    set_caption(caption = 'Studies reporting complete smoking status') %>%
  colformat_double(j = c("Sample size", "Current smokers", "Former smokers", "Never smokers"),
                digits = get_flextable_defaults()$digits, big.mark = get_flextable_defaults()$big.mark, na_str = get_flextable_defaults()$na_str) %>%
  set_table_properties(width = 1, layout = 'autofit')

@@

b <- b %>%
  select(-"Study ID") %>%
  flextable() %>%
    set_caption(caption = 'Studies reporting complete smoking status') %>%
  colformat_double(j = c("Sample size", "Current/former smokers", "Never smokers"),
                digits = get_flextable_defaults()$digits, big.mark = get_flextable_defaults()$big.mark, na_str = get_flextable_defaults()$na_str) %>%
  set_table_properties(width = 1, layout = 'autofit')

@@

numeric_names <- colnames(b[2:9])
b <- b %>%
  select(-"Study ID") %>%
  flextable() %>%
    set_caption(caption = 'Studies reporting complete smoking status') %>%
  colformat_double(j = numeric_names,
                digits = 0, na_str = '-', big.mark = ',') %>%
  set_table_properties(width = 1, layout = 'autofit')

@@

# Smoking prevalence by country {.tabset .tabset-fade .tabset-pills}
``` {r smoking_country, quietly=T}

@@

source(here("scripts", "smoking_plots.R"))
```
We have split the data into a list where each country can be analysed separately if required.

@@

a <- flextable(a) %>%
    set_caption(caption = 'Characteristics of included studies') %>%
  colformat_double(j = numeric_columns, digits = 1, na_str = '-', big.mark = ',') %>%
  colformat_double(j = 'Sample size', digits = 0, na_str = '-', big.mark = ',') %>%
  set_table_properties(width = 1, layout = 'autofit')

@@

numeric_columns <- c('Population_tested', 'SARS_CoV_2_negative', 'N_current_smoker', 'N_former_smoker',
                     'N_current_former_smoker', 'N_never_smoker', 'N_not_stated', 'SARS_CoV_2_positive', 'P_current_smoker',
                     'P_former_smoker', 'P_current_former_smoker', 'P_never_smoker', 'P_not_stated')
a <- flextable(a) %>%
    set_caption(caption = 'SARS-CoV-2 infection by smoking status') %>%
  colformat_double(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',')
```
The number of studies reporting on infection by smoking status is: `r paste(nrow(table_2_word))`
``` {r}
a <- set_header_labels(a,
                       Population_tested = 'Total population tested',
                       SARS_CoV_2_negative = "N (%)",
                       N_current_smoker = "Current smoker (%)",
                       N_former_smoker = "Former smoker (%)",
                       N_current_former_smoker = "Current/former smoker (%)",
                       N_never_smoker = "Never smoker (%)",
                       N_not_stated = "Not stated (%)",
                       SARS_CoV_2_positive = "N (%)",
                       P_current_smoker = "Current smoker (%)",
                       P_former_smoker = "Former smoker (%)",
                       P_current_former_smoker = "Current/former smoker (%)",
                       P_never_smoker = "Never smoker (%)",
                       P_not_stated = "Not stated (%)") %>%
  add_header_row(top = TRUE, values = c("","SARS-CoV-2 negative", "SARS-CoV-2 positive" ), colwidths = c(2, 6, 6)) %>%
  theme_booktabs() %>%
  fix_border_issues() %>%
  set_table_properties(width = 1, layout = 'autofit')  %>%
  colformat_char(na_str = '-')
a

@@

a <- flextable(a) %>%
    set_caption(caption = 'COVID-19 hospitalisation by smoking status') %>%
  colformat_double(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
  colformat_char(na_str = '-')

@@

a <- flextable(a) %>%
    set_caption(caption = 'COVID-19 severity by smoking status') %>%
  colformat_double(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
  colformat_char(na_str = '-')

@@

a <- flextable(a) %>%
    set_caption(caption = 'COVID-19 mortality by smoking status') %>%
  colformat_double(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',') %>%
  colformat_char(na_str = '-')

@@

country_smoking <-  readxl::read_xlsx(here("data_spreadsheet", "data_extraction_current.xlsx"), sheet = 'national_smoking_prevalence') %>%
  flextable()
save_as_docx(country_smoking, path = here('output_data', 'Supplementary_3.docx'))
```
